Robust Estimation, Prediction and Control with Dynamics Structure

### 2020-01-13 main.tex
Updates from Overleaf

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{caption}
\usepackage{blkarray}
\usepackage{todonotes}
\usepackage{mathtools}
\usepackage[backend=biber,style=authoryear,natbib=true,url=false]{biblatex}

@@

\title{An Integrated Framework for Robust Estimation, Prediction and Control}
\author{Edouard Leurent, Denis Efimov, Odalric-Ambrym Maillard}
\date{November 2019}

@@

@@

We also have access to a measurement of the form
$
    \dot{x}(t) + C\eta(t)
$
that we will actually denote (since the control $u(t)$ is known) as:
\begin{equation}
\label{eq:measurement}
    y(t) = \dot{x}(t) - Bu(t) + C\eta(t)
\end{equation}
where $\eta(t)\in\Real^r$ is a measurement noise and $C\in\Real^{p\times r}$ is known. Assumptions over the disturbance $\omega$ and noise $\eta$ will be detailed further.

@@

\paragraph{Objective}% The model estimation procedure is dedicated to building a confidence region around the unknown true model $A(\theta)$. 
Having observed at times $(t_n)_{n\in[N]}$ a history of transitions $\cD = \{(x_n = x(t_n), u_n = u(t_n), y_n = y(t_n))\}_{n\in[N]}$, and given a confidence level $\delta\in[0, 1]$, our goal is to find a confidence region $\cC_\delta$ as tight as possible and such that it holds that:
\begin{align}
\probability{A(\theta)\in \cC_\delta} \geq 1-\delta
\label{eq:polytope}
\end{align}

This involves the trilemma of confidence intervals, where one of three things must be true:
\begin{enumerate}
    \item $A(\theta)$ is in $\cC_\delta$
    \item We got data that was very ($\leq\delta$) unlikely under all possible values $A(\theta)$
    \item Our model is wrong
\end{enumerate}

@@

We further assume that the dynamics $A(\theta)$ enjoy the following additional structure:
\begin{assumption}[Linear Parametrisation]
\label{assumpt:linear_param}
There exists a known tensor $\phi\in \Real^{d \times p \times p}$ such that for all $\theta\in\Theta$:
\begin{equation}
    A(\theta) = \theta^T \phi \eqdef \sum_{i=1}^d \theta_i\phi_i
\end{equation}
where $\phi_i\in\Real^{p\times p}$ for all $i\in[d]$.
\end{assumption}

\begin{assumption}[Noise Model]
\label{assumpt:noise}
At each time $t\geq0$ the measurement noise $\eta(t)$ and disturbance $\omega(t)$  are independent Gaussians with respective means $0_r, 0_s$ and covariances $\Sigma_r, \Sigma_s$:
\begin{equation*}
    \eta(t)\sim \cN(0_r,\Sigma_r),\quad\omega(t)\sim \cN(0_s,\Sigma_s)
\end{equation*}
We further require that the matrix $\Sigma_p = C\Sigma_r C^T + D\Sigma_s D^T \in \Real^{p\times p}$ is full-rank.
\end{assumption}

Note that the full-rank condition in \autoref{assumpt:noise} would be harder to fulfil were it formulated only for the disturbance $D\Sigma_s D^T$, whereas it would be quite reasonable to assume, for instance, that $r=p$, which is sufficient.

\begin{proposition}[Maximum Likelihood Estimate]
\label{prop:mle}
The maximum likelihood estimate for $\theta$ minimises a weighted mean square error:
\begin{equation*}
    \argmax_{\theta\in\Real^d} \probability{\cD\condbar\eqref{eq:dynamics};\Sigma} = \argmin_{\theta\in\Real^d} \sum_{n=1}^N \|y_n -\theta^T\phi x_n\|_{\Sigma_p^{-1}}^2
\end{equation*}
\end{proposition}
\begin{proof}
For $n\in[N]$, at time $t_n$, we have
\begin{align*}
    \probability{y_n, x_n} 
    &= \probability{y_n|x_n}\probability{x_n}\\
    &= \probability{y_n = \dot{x}_n - Bu_n +C\eta_n|x_n,u_n}\probability{x_n}\\
    &= \probability{A(\theta)x_n + B u_n + D \omega_n - Bu_n +C\eta_n = y_n|x_n}\probability{x_n}\\
    &= \probability{C\eta_n + D \omega_n = y_n- \theta^T \phi x_n |x_n}\probability{x_n}\\
    &= \cN(y_n - \theta^T \phi x_n; 0_p,\Sigma_p)\probability{x_n}
\end{align*}
By independence of the noise at different times, we have:
\begin{align*}
    \log\probability{\cD;\Sigma} &= \log\prod_{n=1}^N\probability{y_n, x_n}\\
    &= \sum_{n=1}^N\log \cN(y_n- \theta^T \phi x_n;0_p,\Sigma_p) + \log\probability{x_n}\\
    &= \sum_{n=1}^N -\frac{1}{2} \|y_n -\theta^T\phi x_n\|_{\Sigma_p^{-1}}^2 - \log\left((2\pi)^{p/2} \det(\Sigma_p)^{1/2}\right) + \log\probability{x_n}
\end{align*}
\end{proof}

We stack the observations $y(t_n), x(t_n)$ in $\cD$, as $Y_{[N]}, X_{[N]}\in\Real^{N\times p}$ and the noises $C\eta(t_n)+D\omega(t_n)$ as $\Omega_{[N]} \in\Real^{N\times p}$ to obtain the regression problem:

\begin{equation}
    \label{eq:regression}
    Y_{[N]} = \Phi_{[N]} \theta + \Omega_{[N]},
\end{equation}
with $\Phi_{[N]} = X_{[N]}\phi^T\in \Real^{N\times p\times d}$.

% As detailed in \autoref{prop:mle}, we wish to solve:
% \begin{equation*}
%     \min_{\theta\in\Real^d} \|Y_{[N]} -\Phi_{[N]}\theta\|_{\Sigma_p^{-1}}^2
% \end{equation*}

% where we denote $\|z\|_S^2 = z S z^T$ for $z\in\Real^{N\times p}$ and $S\in\Real^{p\times p}$.

\subsection{Scalar formulation}

\begin{proposition}[Decoupling the noise]
Without loss of generality, up to a change of basis, we can assume that $\Sigma_d$ is diagonal.
\end{proposition}
\begin{proof}
Since $\Sigma_p$ is a symmetric real matrix, it is diagonalisable: there exists $Q\in\cO(p)$ and a diagonal matrix $R$ such that $\Sigma_p = QRQ^{-1}$. Hence, from \eqref{eq:dynamics} and \eqref{eq:measurement}:
\begin{align*}
    Q^{-1}y'(t) &= \sum_{i=1}^d \theta_i Q^{-1}\phi_i x(t) + Q^{-1}(C\eta(t) + D\omega(t))
\end{align*}
By rewriting $y \leftarrow Q^{-1}y, \phi_i\leftarrow Q^{-1}\phi_i, C\leftarrow Q^{-1}C$, and $D\leftarrow Q^{-1}D$ we recover \eqref{eq:dynamics} and \eqref{eq:measurement} in their original form, while now having $C\Sigma_rC^T + D\Sigma_rD^T = R$.
\end{proof}

In order to solve the regression problem \eqref{eq:regression} can simply reshape the two first $N\times p$ dimensions into a single $Np$ dimension. Since $\Sigma_p$ is diagonal, the noises on each row of $\Omega$ are independent, the problem becomes a standard scalar regression problem with heteroscedastic noise (if $\Sigma_p$ is not proportional to $I_p$), and the Lemma 5 of \citet{kirschner18heteroscedastic} can be applied.

\begin{lemma}[Confidence ellipsoid \citep{kirschner18heteroscedastic}]
Assuming that $\Sigma_p$ is diagonal and $\|\theta\|_2\leq S$, it holds with probability $1-\delta$ that:
\begin{equation*}
    \|\theta - \theta_{Np,\lambda}\|_{G_{Np,\lambda}} \leq \sqrt{2\log \frac{\text{det}(G_{Np,\lambda})^{1/2}\text{det}(\lambda I)^{-1/2}}{\delta}} + \lambda^{1/2}S
\end{equation*}
where
\begin{align*}
    \Phi_{[Np]}& \in\Real^{Np\times d},\, Y_{[Np]}\in \Real^{Np} \text{ are the flattened versions of $\Phi_{[N]}$, $Y_{[N]}$};\\
    \tilde{\Phi}_{[Np]} &= \Phi_{[Np]} [\underbrace{\cdots \Sigma_p^{-1} \cdots}_{\text{$N$ times}}]^T, \qquad  \tilde{Y}_{[Np]} = Y_{[Np]} [\underbrace{\cdots \Sigma_p^{-1} \cdots}_{\text{$N$ times}}]^T;\\
    G_{Np, \lambda} &= \tilde{\Phi}^T_{[Np]}\tilde{\Phi}_{[Np]} + \lambda I_d;\\
    \theta_{Np,\lambda} &= G_{Np, \lambda}^{-1} \tilde{\Phi}^T_{[Np]}\tilde{Y}_{[Np]}.
\end{align*}
\end{lemma}

\subsection{Vector formulation}
In the general case, we denote the weighted version of the Frobenius inner product and Frobenius norm as $\inp{M}{N}_{F,\Sigma} = \Tr (M^T \Sigma N)$ and $\|M\|_{F,\Sigma}^2 = \inp{M}{M}_{F,\Sigma}$.

\paragraph{Regularized solution} Since we expect that $\phi x(t)$ may have zero-coordinates during possibly long time intervals (i.e. some features are not always active), we consider instead the weighted $l_2$ regularised problem with parameter $\lambda\in\Real^+_*$ and weights $\Sigma_d\in\Real^{d\times d}$:

\begin{equation}
    \label{eq:regression_min}
    \min_{\theta\in\Real^d} \|(Y_{[N]} -\Phi_{[N]}\theta)^T\|_{F,\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{\Sigma_d}^2
\end{equation}

%For simplicity, we abuse notations and replace $Y_{[N]}$ and $\Phi_{[N]}$ by their transposed forms $Y_{[N]}^T\in\Real^{p\timesN}$ and $\Phi_{[N]}^T\in\Real^{d\times p\times N}$.

@@

\begin{theorem}[Regularized solution] The solution to \eqref{eq:regression_min} is:
\begin{equation}
    \label{eq:vector_rls}
    \theta_{N,\lambda} = G_{N, \lambda}^{-1} \begin{bmatrix}\inp{\Phi_{[N],i}^T}{Y_{[N]}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]}
\end{equation}
where $\Phi_{[N],i}\in\Real^{p\times d}$ is the $i^\text{th}$ component of $\Phi_{[N]}$ along the first dimension, and
\begin{equation*}
    G_{N, \lambda} = \begin{bmatrix}\inp{\Phi_{[N],i}^T}{\Phi_{[N],j}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i,j\in[d]^2} + \lambda \Sigma_d \in \Real^{d\times d}.
\end{equation*}
\end{theorem}

\begin{proof}
In this proof, we drop the $[N]$ subscripts for simplicity.
Let $i\in[d]$ and denote $Y_{-i} = Y - \sum_{j\in[d], j\neq i} \theta_j\Phi_{j}$.

We differentiate the objective $J(\theta) = \|(Y -\Phi\theta)^T\|_{F,\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{\Sigma_d}^2$ as in  \eqref{eq:regression_min} with respect to $\theta_i$:
\begin{align*}
    J(\theta) &= \|(Y - \sum_{j=1}^d\theta_j\Phi_{j})^T\|_{F,\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{\Sigma_d}^2\\
    &=  \|(Y_{-i} - \theta_i\Phi_{i})^T\|_{F,\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{\Sigma_d}^2\\
    &= \Tr\left((Y_{-i} - \theta_i\Phi_{i})\Sigma_p^{-1}(Y_{-i} - \theta_i\Phi_{i})^T\right) + \lambda\|\theta\|_{\Sigma_d}^2\\
    &= \theta_i^2\Tr\left(\Phi_{i}\Sigma_p^{-1}\Phi_{i}^T\right) 
    -\theta_i\Tr\left(\Phi_{i}\Sigma_p^{-1}Y_{-i}^T\right) 
    -\theta_i\Tr\left(Y_{-i}\Sigma_p^{-1}\Phi_{i}^T\right)
    + \Tr\left(Y_{-i}\Sigma_p^{-1}Y_{-i}^T\right)
    \\
    &\quad+ \lambda\|\theta\|_{\Sigma_d}^2
\end{align*}

Hence, 
\begin{align*}
    \nabla_{\theta_i} J(\theta) &= 2\theta_i\Tr\left(\Phi_{i}\Sigma_p^{-1}\Phi_{i}^T\right)
    -2\Tr\left(\Phi_{i}\Sigma_p^{-1}Y_{-i}^T\right) 
    +2\lambda\theta^T\Sigma_{d,i}\\
    &= 2\theta_i\Tr\left(\Phi_{i}\Sigma_p^{-1}\Phi_{i}^T\right)
    +2\sum_{j\neq i}\theta_j\Tr\left(\Phi_{i}\Sigma_p^{-1}\Phi_j^T\right) 
    -2\Tr\left(\Phi_{i}\Sigma_p^{-1}Y^T\right) 
    +2\lambda\theta^T\Sigma_{d,i}\\
    &= 2\theta^T\begin{bmatrix}
    \Tr\left(\Phi_i \Sigma_p^{-1} \Phi_j^T\right)
    \end{bmatrix}_{j\in[d]}
    -2\Tr\left(\Phi_{i}\Sigma_p^{-1}Y^T\right) 
    +2\lambda\theta^T\Sigma_{d,i}\\
\end{align*}

Since, $\nabla_{\theta}J(\theta) = 0 \iff \forall i,\, \nabla_{\theta_i}J(\theta) = 0$, we obtain the following optimality condition:

\begin{align*}
    \forall i,  \theta^T\left(\lambda\Sigma_{d,i} + \begin{bmatrix}
    \Tr\left(\Phi_i \Sigma_p^{-1} \Phi_j^T\right)
    \end{bmatrix}_{j\in[d]}\right)
     = \Tr\left(\Phi_{i}\Sigma_p^{-1}Y^T\right)
\end{align*}
By transposing and arranging the $[d]$ equations in matrix form,
\begin{align*}
    \left(\lambda\Sigma_{d} + \begin{bmatrix}
    \inp{\Phi_{i}^T}{\Phi_{j}^T}_{F,\Sigma_p^{-1}}
    \end{bmatrix}_{i,j\in[d]^2}\right)\theta
     = \begin{bmatrix}\inp{\Phi_{i}^T}{Y^T}_{F,\Sigma_p^{-1}}\end{bmatrix}_{i\in[d]}
\end{align*}

\paragraph{Confidence ellipsoid}

In order to study the concentration of $\theta_{N,\lambda}$, we reinject the expression of $Y$ into \eqref{eq:vector_rls}:
\begin{align*}
    \theta_{N,\lambda} &= G_{N, \lambda}^{-1} \begin{bmatrix}\inp{\Phi_{[N],i}^T}{\sum_{i=j}^d\theta_j\Phi_{[N],j}^T  + \Omega_{[N]}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]}\\
    &= G_{N, \lambda}^{-1}\begin{bmatrix}\inp{\Phi_{[N],i}^T}{\Omega_{[N]}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]} + 
    G_{N, \lambda}^{-1} \sum_{i=j}^d\theta_j\begin{bmatrix}\inp{\Phi_{[N],i}^T}{\Phi_{[N],j}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]}\\
    &= G_{N, \lambda}^{-1}\begin{bmatrix}\inp{\Phi_{[N],i}^T}{\Omega_{[N]}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]} + 
    G_{N, \lambda}^{-1}\left(G_{N, \lambda}-\lambda\Sigma_d\right)\theta\\
    &= \theta + G_{N, \lambda}^{-1}\begin{bmatrix}\inp{\Phi_{[N],i}^T}{\Omega_{[N]}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]} -
    G_{N, \lambda}^{-1}\lambda\Sigma_d\theta
\end{align*}

Hence, for all $x\in\Real^d$,
\begin{align*}
    x^T\theta_{N,\lambda}  -x^T\theta &= x^T G_{N, \lambda}^{-1}\begin{bmatrix}\inp{\Phi_{[N],i}^T}{\Omega_{[N]}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]} 
    - x^T G_{N, \lambda}^{-1}\lambda\Sigma_d\theta\\
    &= \inp{x}{\begin{bmatrix}\inp{\Phi_{[N],i}^T}{\Omega_{[N]}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]}}_{G_{N, \lambda}^{-1}} - \lambda\inp{x}{\Sigma_d\theta}_{G_{N, \lambda}^{-1}}
\end{align*}

Using the Cauchy-Schwartz inequality, we get:
\begin{align*}
    |x^T\theta_{N,\lambda}  -x^T\theta| &\leq \|x\|_{G_{N, \lambda}^{-1}}\left(\left\|\begin{bmatrix}\inp{\Phi_{[N],i}^T}{\Omega_{[N]}^T}_{F, \Sigma_p^{-1}}\end{bmatrix}_{i\in[d]}\right\|_{G_{N, \lambda}^{-1}} + \lambda\|\Sigma_d\theta\|_{G_{N, \lambda}^{-1}}\right)
\end{align*}
\todo[inline]{EL: The first term is a $\Real^d$ vector whose coefficients are inner product of matrices (a sum along both the time $N$ and $p$ dimensions). How to analyze its concentration? For comparison, in \citep{Abbasi2011} $X^T\eta$ was a  $\Real^d$ vector whose coefficients were inner product of vectors $(X_i^T\eta)_i\in[d]$ (a sum over time $N$)}

\paragraph{Confidence polytope}

Given a confidence ellipsoid $C_\delta$, we can enclose it by a polytope $P_\delta$:
\[\cC_\delta \subset \cP_\delta\]
where $\cP_\delta$ can be written as:
\begin{equation}
     \exists A_0,\Delta A_{1:p}: \cP_\delta = \left\{ A_{0}+\sum_{i=1}^{p}\lambda_{i}\Delta A_{i}: \lambda\in[0, 1]^p,  \sum_{i=1}^{p}\lambda_{i}=1\right\}.
\end{equation}

@@

\begin{example*}
Consider a scalar system:
\[
\dot{x}(t)=-\theta(t)x(t)+d(t),\;t\geq0,
\]
where $x(t)\in\Real$ with $x(0)\in[\underline{x}_{0},\overline{x}_{0}]=[1.0, 1.1]$, $\theta(t)\in\Pi=[\underline{\theta},\overline{\theta}]=[0.5,1.5]$ and $d(t)\in[\underline{d},\overline{d}]=[-0.1,0.1]$ for all $t\geq0$. The \autoref{fig:predictor_example} shows the state-interval obtained with \eqref{eq:interval_predictor}.

\begin{figure}
\begin{centering}
\includegraphics[width=\linewidth]{img/interval_predictor}
\par\end{centering}
\caption{Several trajectories for different values of $\theta\in\Theta$ are shown in blue. The result of prediction by \eqref{eq:interval_predictor} is in red: the predictor is stable and produces tight bounds.}
\label{fig:predictor_example}
\end{figure}
\end{example*}

@@

In this section, we still assume that we have a high-confidence polytope $\cP$ that verifies \eqref{eq:polytope} with probability $1-\delta$.

The \textbf{robust control} objective $v^r$:
\todo[inline]{EL: since the interval dynamics will be deterministic, use open-loop policies, ie sequences of actions $\mathbf{u}$ instead of closed-loop policies $\pi$}

\begin{equation}
\label{eq:robust-control}
\max_\pi \underbrace{\min_{\theta \in \Theta} \expectedvalueover{\omega(t)}\left[\sum_{t=0}^\infty \gamma^t r(x(t))\condbar \dot{x}(t) = A(\theta) x(t) + B_u u(t) + B_\omega\omega(t), u_t = \pi(x_t)\right]}_{v^r(\pi)}
\end{equation}

\begin{definition}[Surrogate objective]
We modify the robust objective $v^r$ in \eqref{eq:robust-control} by replacing the true reachability set by its interval prediction $[\underline{x}(t), \overline{x}(t)]$ based on the polytopic approximation $\cP_\delta$ of $A(\Theta)$:

\begin{equation}
\hat{v^r}(\pi) \eqdef \sum_{t=0}^\infty \gamma^t \min_{x\in[\underline{x}(t, \pi), \overline{x}(t, \pi)]}  r(x)
\end{equation}
where $\underline{x}(t, \pi), \overline{x}(t, \pi)$ follow the dynamics defined in \eqref{eq:interval_predictor}.
\end{definition}

@@

\begin{property}[Lower bound]
\label{prop:lower-bound}
With high probability, the surrogate objective $\hat{v^r}$ is a lower bound of the true objective $v^r$:

\begin{equation}
\probability{\forall \pi, \hat{v^r}(\pi) \leq v^r(\pi)} \geq 1-\delta
\end{equation}
\end{property}

@@

The robust objective error $v^r - \hat{v^r}$ stems from two terms: the approximation of the reachable set by the polytopic and interval approximations and the loss of time-dependency between the states within a single trajectory. If both these approximations are tight enough, maximizing the lower bound $\hat{v^r}$ will increase the true objective $v^r$, which is the idea behind Algorithm \ref{algo:irc}.

@@

>>>>

\documentclass{article}
\usepackage[cm]{fullpage}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{caption}
\usepackage{blkarray}
\usepackage{todonotes}
\usepackage{mathtools}
\usepackage[backend=biber,style=authoryear,natbib=true,url=false]{biblatex}

@@

\title{Robust Estimation, Prediction and Control with Dynamics Structure}
\author{Edouard Leurent, Denis Efimov, Odalric-Ambrym Maillard}
\date{November 2019}

@@

\section{Related Work}

For LQ systems, where parameters A,B are unknown and no structure is available.
\begin{itemize}
    \item OFU: \citep{abbasi-yadkori11a}. Summary: $\theta=(A~B)$, ellipsoid in Frobenius norm + controllability constraint, choose and stabilize (Ricatti) an optimistic candidate, cumulative regret in $O(\sqrt{T}\log\frac{1}{\delta})$.
    \item PSRL with TS: \citep{abeille18a}. Summary: RLS estimate + normal perturbation in the ellipsoid + rejection sampling over a controllability constraint, with frequent update, cumulative regret in $\tilde{O}(\sqrt{T})$
    \item Offline estimation and control synthesis: \citep{Dean2017}. Summary: Estimation procedure of $N$ episodes of duration $T$ with gaussian excitation, ellipsoid in L2 norm, robust controller with System Level Synthesis, simple regret in $O(\sqrt{\frac{\log(1/\delta)}{N}})$
    \item \citep{Dean2018}
\end{itemize}

Differences with our work:
\begin{itemize}
    \item We with to act in a reasonable (structure) but cautious (robust) way, and progressively get more and more aggressive as we obtain more data and reduce uncertainty.
    \item Thus, we do not wish to explore actively the more promising regions as in OFU, nor the more likely region as in TS, nor act randomly as in Coarse-ID.
    \item We have linear dynamics but our cost is not quadratic (to accurately represent planning task with discrete/branching decisions: e.g. take or give way)
    \item Our optimisation problem amounts to discrete controller selection, instead of continuous controller synthesis

\end{itemize}

@@

We also have access to a measurement of the form
$
    \dot{x}(t) + C\nu(t)
$
that we will actually denote (since the control $u(t)$ is known) as:
\begin{equation}
\label{eq:measurement}
    y(t) = \dot{x}(t) - Bu(t) + C\nu(t)
\end{equation}
where $\nu(t)\in\Real^r$ is a measurement noise and $C\in\Real^{p\times r}$ is known. Assumptions over the disturbance $\omega$ and noise $\nu$ will be detailed further.

@@

\paragraph{Objective}
Having observed at times $(t_n)_{n\in[N]}$ a history of transitions $\cD = \{(x_n = x(t_n), u_n = u(t_n), y_n = y(t_n))\}_{n\in[N]}$, and given a confidence level $\delta\in[0, 1]$, our goal is to find a confidence region $\cC_\delta$ as tight as possible and such that it holds that:
\begin{align}
\probability{A(\theta)\in \cC_\delta} \geq 1-\delta
\label{eq:polytope}
\end{align}

% This involves the trilemma of confidence intervals, where one of three things must be true:
% \begin{enumerate}
%     \item $A(\theta)$ is in $\cC_\delta$
%     \item We got data that was very ($\leq\delta$) unlikely under all possible values $A(\theta)$
%     \item Our model is wrong
% \end{enumerate}

@@

We further assume that the dynamics $A(\theta)$ enjoy the following additional structure:
\begin{assumption}[Linear parametrisation]
\label{assumpt:linear_param}
There exists a known tensor $\phi\in \Real^{d \times p \times p}$ such that for all $\theta\in\Theta$:
\begin{equation}
    A(\theta) = \theta^\transp \phi \eqdef \sum_{i=1}^d \theta_i\phi_i
\end{equation}
where $\phi_i\in\Real^{p\times p}$ for all $i\in[d]$.

For all $n\geq 0$, we also denote $\Phi_n = [\phi_1 x_n \cdots \phi_d x_n]\in\Real^{p\times d}$.
\end{assumption}

\paragraph{Regularised least square} Since we expect that $\Phi_n$ may have zero-coordinates during possibly long time intervals (i.e. some features are not always active), we consider the weighted $L_2$-regularised regression problem with weights  $\Sigma_p\in\Real^{p\times p}$ and $\Sigma_d\in\Real^{d\times d}$ and parameter $\lambda\in\Real^+_*$:

\begin{equation}
    \label{eq:regression_min}
    \min_{\theta\in\Real^d} \sum_{n=1}^N \|y_n -\Phi_n\theta\|_{\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{\Sigma_d}^2
\end{equation}

@@

\begin{theorem}[Regularised solution]
\label{thm:regularized_solution}
The solution to \eqref{eq:regression_min} is:
\begin{equation}
    \label{eq:vector_rls}
    \theta_{N,\lambda} = G_{N, \lambda}^{-1} \sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1} y_n
\end{equation}
where 
\begin{equation*}
    G_{N, \lambda} = \sum_{n=1}^N \Phi_{n}^\transp\Sigma_p^{-1}\Phi_{n}  + \lambda \Sigma_d \in \Real^{d\times d}.
\end{equation*}
\end{theorem}

Denote $\eta(t) = C\nu(t) + D\omega(t)$. By substituting the expression of $y_n$ into \eqref{eq:vector_rls}, we obtain:
\begin{lemma}[Regression error]
\label{lem:regression-error}
\begin{align*}
    \theta_{N,\lambda} - \theta = G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n - G_{N, \lambda}^{-1}\lambda\Sigma_d\theta 
\end{align*}
\end{lemma}

Depending on the assumption we take over the noise $\eta_n$ and the true parameter $\theta$, we can bound the regression error $\theta_{N,\lambda} - \theta$ in different ways.

\subsection{Sub-Gaussian noise}

We say that a random vector $x\in\Real^k$ is sub-Gaussian with covariance proxy $\Sigma$, and denote $x\sim Sub_G(\Sigma)$, if $\expectedvalue[x] = 0_k$ and:
\[
\forall u\in\Real^k, \expectedvalue \left[ \exp{\left( u^\transp x\right)}\right] \leq \exp{\left( \frac{1}{2} u^\transp \Sigma u\right)}
\]

\begin{assumption}[Sub-Gaussian Noise Model]
\label{assumpt:noise}
At each time $t\geq0$ the sum $\eta(t)$ of projected measurement noise $\nu(t)$ and disturbance $\omega(t)$ is a Sub-Gaussian noises with covariance proxy $\Sigma_p \in \Real^{p\times p}$:
\begin{equation*}
    \eta(t) = C\nu(t) + D\omega(t) \sim Sub_G(\Sigma_p)
\end{equation*}

For instance, this is the case of $\nu$ and $\omega$ are Gaussian noises with respective covariance matrices $\Sigma_r$ and $\Sigma_s$, with $\Sigma_p \eqdef C\Sigma_r C^\transp + D\Sigma_s D^\transp$.
\end{assumption}

\begin{lemma}[Theorem 1 of \citep{Abbasi2011}, Matrix version]
\label{lem:concentration}
Let $\{F_t\}_{t=0}$ be a filtration.
Let $\{\eta_t\}_{t=1}^\infty$ be a $\Real^p$-valued stochastic process such that $\eta_t$ is $F_t$-measurable and conditionally $\Sigma_p$-SubGaussian.

Let $\{\Phi_t\}_{t=1}^\infty$ be an $\Real^{p\times d}$-valued stochastic process such that $\Phi_t$ is $F_t$-measurable. Assume that $G$ is a $d\times d$ positive definite matrix. For any $t\geq 0$, define:

\begin{equation*}
    \overline{G}_t = G + \sum_{s=1}^t \Phi_s^\transp \Sigma_p^{-1} \Phi_s \in \Real^{d\times d} \qquad S_t = \sum_{s=1}^t \Phi_s^\transp\Sigma_p^{-1}\eta_s \in \Real^{d}
\end{equation*}
Then, for any $\delta>0$, with probability at least $1-\delta$, for all $t\geq0$

\begin{align*}
    \|S_t\|_{\overline{G_t}} \leq \cdots
\end{align*}
\end{lemma}

\begin{theorem}[Confidence ellipsoid]
\label{thm:confidence_ellipsoid}
\todo[inline]{TODO}
\end{theorem}

\begin{lemma}[Confidence polytope]
\label{lem:confidence_polytope}
We can enclose the confidence ellipsoid $C_\delta$ obtained in $\eqref{eq:confidence-ellipsoid}$ within a polytope $P_\delta$:
\[\cC_\delta \subset \cP_\delta\]
where $\cP_\delta$ can be written as:
\begin{equation}
     \cP_\delta = \left\{ A_{0}+\sum_{i=1}^{2^d}\lambda_{i}\Delta A_{i}: \lambda\in[0, 1]^{2^d},  \sum_{i=1}^{2^d}\lambda_{i}=1\right\}.
\end{equation}
with 
\begin{align*}
    &G_{N,\lambda}^{-1} = PDP^{-1}\\
    &h_k \text{ is the }k^\text{th}\text{ element of }\{-1,1\}^d\text{, for } k\in[2^d]\\
    &\Delta\theta_k = C_{N,\lambda,\delta}^{1/2} P^{-1}D^{-1/2} h_k \\
    &A_0 = \theta_{N,\lambda}^\transp\Phi\\
    &\Delta A_k = \Delta\theta_k^\transp\Phi\\
\end{align*}
\end{lemma}

\begin{figure}
    \centering
    \includegraphics[trim={3.8cm, 2cm, 5cm, 3.8cm}, clip, width=0.6\textwidth]{img/ellipsoid_to_polytope}
    \caption{From the confidence ellipsoid $\cC_\delta$ to its enclosing polytope $\cP_\delta$}
    \label{fig:ellipsoid_to_polytope}
\end{figure}

\subsection{Bounded noise}

\begin{assumption}[Bounded noise]
The noise $\eta(t) = C\nu(t) + D\omega(t)$ is bounded.
\[\|\eta\|_\infty < \infty\]
\end{assumption}

Then, 
\[
\|\theta_{N,\lambda} - \theta\|_1 \leq \|G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\|_1\|\eta\|_\infty + \lambda\|G_{N, \lambda}^{-1}\Sigma_d\theta\|_1
\]

This $L_1$-ball is a polytope with $2d$ vertices, that we denote $\cP$.

@@

% \begin{example*}
% Consider a scalar system:
% \[
% \dot{x}(t)=-\theta(t)x(t)+d(t),\;t\geq0,
% \]
% where $x(t)\in\Real$ with $x(0)\in[\underline{x}_{0},\overline{x}_{0}]=[1.0, 1.1]$, $\theta(t)\in\Pi=[\underline{\theta},\overline{\theta}]=[0.5,1.5]$ and $d(t)\in[\underline{d},\overline{d}]=[-0.1,0.1]$ for all $t\geq0$. The \autoref{fig:predictor_example} shows the state-interval obtained with \eqref{eq:interval_predictor}.

% \begin{figure}
% \begin{centering}
% \includegraphics[width=\linewidth]{img/interval_predictor}
% \par\end{centering}
% \caption{Several trajectories for different values of $\theta\in\Theta$ are shown in blue. The result of prediction by \eqref{eq:interval_predictor} is in red: the predictor is stable and produces tight bounds.}
% \label{fig:predictor_example}
% \end{figure}
% \end{example*}

@@

We consider the robust control objective $v^r$:
\todo[inline]{EL: since the interval dynamics will be deterministic, use open-loop policies, ie sequences of actions $\mathbf{u}$ instead of closed-loop policies $\pi$}

\begin{equation}
\label{eq:robust-control}
\max_\pi \underbrace{\min_{A(\theta) \in \cP} \left[\sum_{t=0}^\infty \gamma^t r(x(t))\condbar u_t = \pi(x_t), \omega(t)\right]}_{v^r(\pi)}
\end{equation}

\begin{definition}[Surrogate objective]
We approximate the robust objective $v^r$ in \eqref{eq:robust-control} by replacing the true reachability set by its interval prediction $[\underline{x}(t), \overline{x}(t)]$ based on the polytopic approximation $\cP_\delta$ of $A(\Theta)$:

\begin{equation}
\hat{v}^r(\pi) \eqdef \sum_{t=0}^\infty \gamma^t \min_{x\in[\underline{x}(t, \pi), \overline{x}(t, \pi)]}  r(x)
\end{equation}
where $\underline{x}(t, \pi), \overline{x}(t, \pi)$ follow the dynamics defined in \eqref{eq:interval_predictor}.
\end{definition}

@@

\begin{property}[Lower bound]
\label{prop:lower-bound}
With high probability, the surrogate objective $\hat{v}^r$ is a lower bound of the true objective $v^r$:

\begin{equation}
\hat{v}^r(\pi) \leq v^r(\pi)
\end{equation}

If we assume that $R$ is $L$-Lipschitz, then we can bound the other side:
\begin{equation*}
     \hat{v}^r(\pi) \leq v^r(\pi) \leq \hat{v}^r(\pi) + L\sum_{t=0}^\infty \gamma^t \mu([\underline{x}_N(t),\overline{x}_N(t)])
\end{equation*}
\end{property}

@@

The robust objective error $v^r - \hat{v^r}$ stems from two terms: the approximation of the reachable set by the polytopic and interval approximations and the loss of time-dependency between the states within a single trajectory. If both these approximations are tight enough, maximising the lower bound $\hat{v^r}$ will increase the true objective $v^r$, which is the idea behind Algorithm \ref{algo:irc}.

@@

\appendix

\section{Proofs}

\subsection{Proof of \autoref{thm:regularized_solution}}
